---
title: "RedRibbon: a rank-rank hypergeometric overlap libray"
author: "Anthony Piron"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{RedRibbon}
  %\VignetteEngine{knitr::knitr}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(kableExtra)
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Synthetic dataset


First, we prepare a synthetic dataset in a data.frame `df` for demonstration purpose. RedRibbon will be applied to this dataset in this vignette.

The column `a` and `b` are mandatory. Those column are the two lists over which the rank-rank hyper-geometric is run.

```{r dataset, results='hide'}
n <- 1000
n2 <- n %/% 2
n4 <- n %/% 4
a <- (1:n) - n2
b <- a
a[n4:(n4+n2-1)] <- rnorm(n2, sd=100)
b[n4:(n4+n2-1)] <- rnorm(n2, sd=100)
df <- data.frame(
	id = paste0("gene", 1:n),
	a = a,
	b = b)
```

`df` contains half features perfectly matching perfectly between `a` and `b` and the remaining elements are random, `head(df)` gives
	
```{r df, echo=FALSE}
head(df) %>%
  kbl() %>%
  kable_styling()
```

# Synopsis


Load the library with

```{r Library loading}
library(RedRibbon)
```

Create an S3 RedRibbon object,

```{r s3, results='hide'}
rr <- RedRibbon(df, enrichment_mode="hyper-two-tailed")
```

Compute the enrichment and adjusted p-value with the `permutation` option set to compute the adjusted p-value and `whole` to split the overlap map into four quadrants (down-down, up-up, down-up, up-down),

```{r enrichment, results='hide'}
quad <- quadrants(rr, algorithm="ea", permutation=TRUE, whole=FALSE)
```

# Plots

Plots the results,

```{r ggplot}
gg <- ggplot(rr, quadrants=quad) + coord_fixed(ratio = 1)
gg
```

The `gg` variable contains a standard ggplot2 object,
```{r class ggplot}
class(gg)
```

It can be manipulated a such, for example, with `+ coord_fixed(ratio = 1)` to fix the aspect ratio of the plot.


# Enrichment statistics

The list of genes enriched in the down-down direction given with `df[quad$downdown$positions,]`. Here are the 6 first rows,
```{r genes, echo=FALSE}
head(df[quad$downdown$positions,])  %>%
  kbl() %>%
  kable_styling()
```

The following statistics are available,

```{r names}
names(quad$downdown)
```

`(i , j)` is the  minimal p-value coordinate on the enrichment map, `pvalue` and `log_pvalue` are the minimal p-value, `direction` is the direction of the enrichment (same direction is postive -- up-up, down-down --, opposite is negative -- down-up, up-down) and `count` is the number of genes overlapping.

# Compatibility mode


This mode only exists for compatibility and is obsolete. Please use the new interface shown here above.

```{r compat}
compat <- RRHO(df[,c("id", "a")], df[,c("id", "b")], plots=TRUE, stepsize=n %/% 10)
compat$gg
```

```{r hypermat, results='hide'}
compat$hypermat
```
```{r hypermat2, echo=FALSE}
compat$hypermat %>%
  kbl() %>%
  kable_styling()
```



